Dysregulation of Cerebral Hemoglobin Content in Malawian Children with Malaria

### 2021-08-20 figures.Rmd
ROC curves

load("glm_model.Rdata")
load("results.Rdata")
load("dfa_example.Rdata")
load("dfa_results.Rdata")

@@

>>>>

load("glm_model.Rdata")
load("results.Rdata")
load("roc_curves.Rdata")
load("dfa_example.Rdata")
load("dfa_results.Rdata")

@@

#### C. ROC curve of final model compared to other iterations

```{r fig3c_prep, eval = FALSE}

## write function to create model on given df

f_log_reg_model <- function(df_model) {

  # resample
  hbox_boot <- bootstraps(df_model, strata = Status, times = 1000)

  # recipe
  hbox_rec <- recipe(Status ~ ., data = df_model) %>% 
    step_BoxCox(all_predictors()) %>% 
    step_nzv(all_predictors()) %>% 
    step_smote(Status) # upsample to deal with unbalanced sample sizes

  # create model on resamples
  glm_res <- hbox_wf %>% 
    add_model(glm_spec) %>%
    fit_resamples(
      resamples = hbox_boot,
      metrics = metric_set(roc_auc, accuracy, sensitivity, specificity),
      control = tune::control_resamples(save_pred = TRUE, verbose = TRUE)
    )

  # finalize model
  hbox_final <- hbox_wf %>% 
    add_model(glm_spec) %>% 
    finalize_workflow(glm_best) %>%
    fit(df_model) %>%
    pull_workflow_fit()

  # view coefficients & p-values
  model_coefs <- hbox_final %>% tidy()

  # return list including model and a table of best model variable statistics
  list(glm_res, model_coefs)

## write function to collect predictions for ROC plot

f_collect_preds <- function(glm_res) {

      glm_res %>% 
      collect_predictions() %>% 
      roc_curve(Status, .pred_CM)

## Perform functions on all iterations of model

set.seed(123)

    # select starting variables
    df_hbox_model1 <- df_hbox_log_reg %>% 
      dplyr::select(c(Status, Hematocrit, Lactate, avg_Hb_o2sat, Hb_o2sat_second_alpha, avg_Hb_conc, Hb_conc_second_alpha)) %>% 
      mutate(Status = factor(Status, levels = c("CM", "UM")))

    # model 1: Status ~ Hematocrit + Lactate + avg_Hb_o2sat + Hb_o2sat_second_alpha + avg_Hb_conc + Hb_conc_second_alpha
    model1 <- f_log_reg_model(df_hbox_model1)
    roc1 <- f_collect_preds(model1[[1]]) %>% mutate(model = "1")
    model1[[2]] # remove Hb_o2sat_second_alpha (highest p-val)

    # model 2: Status ~ Hematocrit + Lactate + avg_Hb_o2sat + avg_Hb_conc + Hb_conc_second_alpha
    df_hbox_model2 <-  df_hbox_model1 %>% dplyr::select(-Hb_o2sat_second_alpha)
    model2 <- f_log_reg_model(df_hbox_model2)
    roc2 <- f_collect_preds(model2[[1]]) %>% mutate(model = "2")
    model2[[2]] # remove Lactate

    # model 3: Status ~ Hematocrit + avg_Hb_o2sat + avg_Hb_conc + Hb_conc_second_alpha
    df_hbox_model3 <-  df_hbox_model2 %>% dplyr::select(-Lactate)
    model3 <- f_log_reg_model(df_hbox_model3)
    roc3 <- f_collect_preds(model3[[1]]) %>% mutate(model = "3")
    model3[[2]] # remove avg_Hb_o2sat

    # model 4: Status ~ Hematocrit + avg_Hb_conc + Hb_conc_second_alpha
    df_hbox_model4 <-  df_hbox_model3 %>% dplyr::select(-avg_Hb_o2sat)
    model4 <- f_log_reg_model(df_hbox_model4)
    roc4 <- f_collect_preds(model4[[1]]) %>% mutate(model = "4")
    model4[[2]] # none to remove, final model

save(roc1, roc2, roc3, file = "roc_curves.Rdata")

# calculate areas under curves

library(MESS)
roc1 <- roc1 %>% mutate(false_pos = 1 - specificity)
auc(roc1$false_pos, roc1$sensitivity)

roc2 <- roc2 %>% mutate(false_pos = 1 - specificity)
auc(roc2$false_pos, roc2$sensitivity)

roc3 <- roc3 %>% mutate(false_pos = 1 - specificity)
auc(roc3$false_pos, roc3$sensitivity)

roc_final <- glm_res %>% 
  collect_predictions() %>% 
  roc_curve(Status, .pred_0) %>% 
  mutate(false_pos = 1 - specificity)
auc(roc_final$false_pos, roc_final$sensitivity)

```{r fig3c_plot, fig.height = 8}

df_roc <- glm_res %>% 
  collect_predictions() %>% 
  roc_curve(Status, .pred_0) %>% 
  mutate(model = "final") %>% 
  bind_rows(roc1, roc2, roc3)

ggplot() +
  geom_path(data = df_roc %>% filter(model == "final"), 
            aes(x = 1 - specificity, y = sensitivity, color = model), 
            size = 2.5) +
  geom_path(data = df_roc %>% filter(model != "final"), 
            aes(x = 1 - specificity, y = sensitivity, color = model), 
            size = 1) +
  geom_abline(lty = 3) +
  labs(x = "False positive rate", y = "True positive rate") +
  theme_bw() +

  theme(text = element_text(family = "Arial"),
        axis.title = element_text(size = 25),
        axis.text = element_text(size = 20),
        legend.title = element_text(size = 20),
        legend.text = element_text(size = 15),
        legend.position = "top")
